---
title: "Lab 01 - Hello R"
author: "Tiffani Hill"
date: "01/14/2026"
output: 
  html_document: 
    highlight: espresso
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1

There are 1846 rows and 3 columns. The variables are dataset, x, and y.

### Exercise 2

First, let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Next, let's calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```
There is no linear relationship between x and y which is reflected graphically and in the correlation coefficient. 


### Exercise 3

```{r plot-star, fig.height=3, fig.width=4}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Graphically, there seems to be no linear relationship between x and y.

```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```
There isn't a linear relationship between x and y. Thus, the correlation coefficient is very small.


### Exercise 4

```{r-datasaurus-facet}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")
```

```{r-cor_datasaurus}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```

### Exercise 5
```{r plot-bullseye, fig.height=3, fig.width=3}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```
Here, the graphic helps us see the outliers in the circle dataset.